\chapter{Sprint Focus: Git and Navigation Workflow}
\minitoc

\section{Sprint objective}
This sprint improves day-to-day developer flow around branch awareness, push safety, and project navigation.

\section{Git workflow buttons and layout}
UI actions are separated into:
\begin{itemize}
  \item \textbf{Top toolbar}: Open TeX, Save, Build.
  \item \textbf{Git panel}: current branch label, branch table (name/type/date), and actions: Checkout, Pull, Status, Auto Push, Delete Local Branch, Delete Remote Branch, Ignore Pattern.
\end{itemize}
This avoids scattered actions and keeps Git tasks grouped.

\section{Branch ordering by recent activity}
Branch lists are intentionally \textbf{not alphabetical}. Branches are loaded with commit-time sorting so the most recently modified branches are shown first.

\section{Git Ignore Manager}
The panel now includes \textbf{Ignore Pattern}. This dialog supports patterns such as \texttt{*.pyc}, \texttt{\_\_pycache\_\_/}, \texttt{*.aux}, \texttt{*.log}, \texttt{*.fdb\_latexmk}, \texttt{*.fls}, and \texttt{*.xdv}.
\begin{enumerate}
  \item detect Git repository root with \texttt{git rev-parse --show-toplevel},
  \item list tracked files via \texttt{git ls-files},
  \item preview tracked files that match the entered pattern,
  \item append pattern to root \texttt{.gitignore} without duplicates (create file if missing),
  \item optionally run \texttt{git rm --cached -- <files...>} to remove matched tracked files from the Git index only.
\end{enumerate}
Important: this does not delete files from disk; it only stops tracking matching generated/cache files.

\section{Branch management safety rules}
\textbf{Local branch deletion}
\begin{itemize}
  \item safe delete uses \texttt{git branch -d <branch>},
  \item deleting the currently checked out branch is blocked,
  \item if safe delete fails (typically unmerged commits), the UI offers explicit force delete with \texttt{git branch -D <branch>} and a warning.
\end{itemize}

\textbf{Remote branch deletion}
\begin{itemize}
  \item remote delete uses \texttt{git push origin --delete <branch>},
  \item protected branches are blocked by default: \texttt{origin/main}, \texttt{origin/master}, \texttt{origin/develop},
  \item confirmation explicitly shows which remote branch will be deleted,
  \item branch refs are refreshed with \texttt{git fetch --prune}.
\end{itemize}

\textbf{Local vs remote deletion}
\begin{itemize}
  \item deleting a local branch affects only your clone,
  \item deleting a remote branch removes it from \texttt{origin} for collaborators.
\end{itemize}

\section{Graphical Auto Push}
Auto Push reimplements the repository \texttt{autopush} behavior with a confirmation dialog:
\begin{enumerate}
  \item detect changes with \texttt{git status --short --branch},
  \item show changed files before commit,
  \item require non-empty commit message,
  \item run \texttt{git add -A}, one \texttt{git commit -m}, then \texttt{git push},
  \item stream command output into the log panel.
\end{enumerate}
If there are no changes, Auto Push exits gracefully with an informational message.

\section{Project explorer / solution explorer}
A left sidebar presents the active project tree and opens \texttt{.tex} files on click. Build artefacts such as \texttt{.aux}, \texttt{.log}, \texttt{.fls}, \texttt{.toc}, \texttt{.out}, \texttt{.xdv}, and \texttt{.mtc*} are hidden by default.

\section{Intentionally postponed}
\begin{itemize}
  \item embedded PDF preview pane,
  \item PTY-grade terminal emulation,
  \item embedded Neovim editing,
  \item in-editor Copilot completions.
\end{itemize}

\section{Future Neovim and Copilot integration}
Planned direction:
\begin{enumerate}
  \item replace or augment the plain text editor with embedded Neovim,
  \item connect Copilot-compatible Neovim plugins,
  \item keep build/Git/project explorer controls in PySide6,
  \item preserve branch-aware status and Auto Push safety checks.
\end{enumerate}
